---
#title: "GaussianNoises and Delta_LL of model comparison project"
#author: "e guven"
#date: "May 3, 2016"
#output: html_document


#EG started this on 20160403 
#how does the noise effect the loglikelihood 
#Weibull model->lambda is a scale, v is a shape parameter >0
#Gompertz Model-> beta(G) is a scale, alpha(R) is a shape parameter>0 

#difference function of loglikelihood function of gompertz and weibull p.d.fs
#test if L(Weibull,X)>L(Gompertz,X) for parameters Weibull model->theta is a scale, lambda is a shape parameter >0
#Gompertz Model-> beta is a scale, alpha is a shape parameter>0 

#For additive Gaussian noise e ~ N (0, sigma^2) with known variance sigma^2
#sd of gaussian noise function
#max sd would be = 3*mean(inverse.gomp.CDF)
#min sd would be mean(inverse.gomp.CDF)



#fix the paramaters of weibull model
---

```{r}
require(flexsurv)
require(gplots)
```


For Gaussian noises X, Likelihood(X, Gompertz) ?<? Likelihood(X, Weibull) 
#theta=0.025
#gamma=0.001
#test G and R in nested for loops
```{r}
beta= 0.034  #G G= (0.1,0.25)
alpha=0.01 #R  R= (0.001,0.1)

N=100 # population size

#generate gompertz random numbers by using inverse CDF
#prediction
rgompertz = function(R,G, n){
  x.uniform = runif(n)
  inverse.gomp.CDF = function(R,G,y) {  (1/G)*log(1 - (G/R)*log(1-y)  ) }
  x.gompertz = inverse.gomp.CDF(alpha,beta, x.uniform)
  return(x.gompertz)
}


rweibull= function(lambda,v,n)
{
  x.uniform= runif(n)
  inverse.wei.CDF=function(lambda,v,y) { lambda*(-log(1-y))^(1/v)}
  x.weibull=inverse.wei.CDF(lambda,v,x.uniform)
  return(x.weibull)
}

#create a function that calculates noise of lifespan
calculate.noise = function(i){
  
  #lifespan
  gaussian<-rnorm(N, mean = 0, sd=i)
  #observation
  #X<- gompertz.random +gaussian to be used in simulation later.
  #noise
  noise=sd(gaussian)
  
  return(noise)
}




#generate gompertz random numbers (lifespan) by using inverse CDF
#prediction
gompertz.random<-rgompertz(alpha,beta,N)
average.lifespan=mean(gompertz.random)

#put the noise values from loops into noise and delta.LL in a list
sderr<- list()
Delta_LL<-list() 
G<-list()
R<-list()
LWei<-list()
LGomp<-list()
MeanLF<-list()
sdLS<-list()
Delta_LL.flex<-list()
LWei.flex<-list()
LGomp.flex<-list()
G.flex.estimated<-list()
R.flex.estimated<-list()
LLG.par<-list()
LLR.par<-list()
v.flex.estimated<-list()
lambda.flex.estimated<-list()

```

```{r,warning=FALSE}
for (beta in c(0.05,0.08, 0.1,0.15,0.17, 0.2, 0.25)){
  for (alpha in c(1E-3, 0.002, 0.005,0.008, 0.01,0.03, 0.05)){ #fix alpha or in other words R shape parameter
    #for (sd in seq(round.lifespan,3*round.lifespan,by=1)){
    for (i in c(0, 0.5, 1,2,3,4, 5)){ 
    #for (i in c(0)){ 
      
      #generate gompertz random numbers (lifespan) by using inverse CDF
      #prediction
      gompertz.random<-rgompertz(alpha,beta,N)
      average.lifespan=mean(gompertz.random)
    
      MeanLF[[length(MeanLF)+1]]=average.lifespan
      
      sd.gaussian=calculate.noise(i)
      gaussian<-rnorm(N, mean = 0, sd=sd.gaussian)
      #standard deviation of lifespan.
      sd.lifespan=sd(gompertz.random)
      sdLS[[length(sdLS)+1]] =sd.lifespan
      
      
      #simulate lifespan plus normal noises
      #observation
      #gompertz.random<-rgompertz(alpha,beta,100)
      #x<- gompertz.random + gaussian
      lifespan<- gompertz.random +gaussian
      
      #Log likelihood function for the Weibull model
      weib.likl<-function(param,y){
        theta<-param[1]
        gamma<-param[2]
        
        y=lifespan[!is.na(lifespan)]
        logl<-sum((log(gamma) + gamma*log(theta) + (gamma-1)*log(y) -
                           (theta*y)^gamma ))
        return(-logl)
      }
      weib=optim(c(3,0.01),weib.likl,y=lifespan,hessian=T)
      
      weib$value
      LWei[[length(LWei)+1]] = weib$value
      
      #Log likelihood of gompertz function
      LL_gomp <- function (param,y){
        
          beta<-param[1]
          alpha<-param[2]
          delta=1
          y=lifespan[!is.na(lifespan)]
          logl<-sum(delta*(log(beta)+alpha*y+(-(beta/alpha)*(exp(alpha*y)-1)))) +
            sum((1-delta)*(-(beta/alpha)*(exp(alpha*y)-1)))
          return(-logl)
        }
        gomp<-optim(c(0.03,0.01),LL_gomp,y=lifespan)
        
      gomp$value
      LGomp[[length(LGomp)+1]] = gomp$value
      
      # store R and G estimation from optim of ll functions in Gompertz
      LLG.par[[length(LLG.par)+1]] =gomp$par[1]
      LLR.par[[length(LLR.par)+1]]=gomp$par[2]
      
      delta.likelihood.wei<-(weib$value-gomp$value)
      
      #calculate LL and noise change
      sderr[[length(sderr)+1]] = i       
      Delta_LL[[length(Delta_LL)+1]] = delta.likelihood.wei
      G[[length(G)+1]]=beta
      #switch to alpha.seq when for fixed beta
      R[[length(R)+1]]=alpha
      
      #todo use flexsurv to calculate the LL
      
      #flexsurv only works with positive variables.
      #fix gaussian std to 0
      
      gaussian.flex= rnorm(N, mean = 0, sd=0)
      X.flex= gompertz.random +gaussian.flex
      
      #test with flexsurv your loglikelihood estimations above
      fitGomp = flexsurvreg(formula = Surv(X.flex) ~ 1, dist="gompertz")
      fitWei = flexsurvreg(formula = Surv(X.flex) ~ 1, dist="weibull")
      
      
      LWei.flex[[length(LWei.flex)+1]]=fitWei$loglik
      
      LGomp.flex[[length(LGomp.flex)+1]]=fitGomp$loglik
      
      param.Gomp<-fitGomp$res; R.flex<-param.Gomp[1]; G.flex<-param.Gomp[2];
      
      R.flex.estimated[[length(R.flex.estimated)+1]]<-R.flex
      G.flex.estimated[[length(G.flex.estimated)+1]]<-G.flex
      
      param.Wei<-fitWei$res; v.flex<-param.Wei[1]; lambda.flex<-param.Wei[2];
      
      v.flex.estimated[[length(v.flex.estimated)+1]]<-v.flex; 
      lambda.flex.estimated[[length(lambda.flex.estimated)+1]]<-lambda.flex
      
      
      
      
      delta_flexsurv=fitWei$loglik-fitGomp$loglik 
      
      #fitWei$loglik
      
      Delta_LL.flex[[length(Delta_LL.flex)+1]]=delta_flexsurv
      
    }
  }
}

```
```{r}
results = data.frame(cbind(sderr), cbind(R),cbind(LLR.par),cbind(R.flex.estimated),cbind(G),cbind(LLG.par),cbind(G.flex.estimated),cbind(Delta_LL) , cbind(Delta_LL.flex),
                     cbind(LWei),cbind(LWei.flex), cbind(LGomp),cbind(LGomp.flex), cbind(MeanLF), cbind(sdLS))

results


results_mat<-as.matrix(results)








