---
title: "Weibull vs Gompertz model comparison with added normal noise"
author: "e guven"
date: "September, 2016"
output: html_document
---

```{r}
library('varhandle')
library('flexsurv')
library('stringr')

```

##load original data for analyzing

```{r}
new_rls<-read.csv("rls.csv")


my.data=new_rls

#my.data<-my.data[complete.cases(my.data$ref_lifespan_mean),]

# uniquely determine the ref_lifespans of each experiments individually
big_data<-my.data[!duplicated(my.data$ref_lifespans),] 


```{r}

#big_data = big_data[complete.cases(big_data$ref_lifespan_mean),]
ref_lifespan_mean<-list()



for (k in 1:(length(sort(unique(big_data$ref_lifespans))))){
#k=94
f1<-big_data$ref_lifespans
f<-unfactor(f1[k])
ref_lifespan_single<-as.numeric(unlist(str_split(f, ",")))
ref_lifespan_single_mean<-mean(ref_lifespan_single)
ref_lifespan_mean[[length(ref_lifespan_mean)+1]]<-ref_lifespan_single_mean

#avg.ref.ls<-round(mean(c(ref_lifespan_single)))
#x<-big_data$id
#ref_lifespan_mean<-round(big_data$ref_lifespan_mean[k])


}

```

```{r}
big_data$ref_lifespan_mean<-unlist(ref_lifespan_mean)



fit_names = c( 'genotype','media','mat','sd.ls','medianLS','Delta_LL','AvgLS','gompLogLik','gompRate','gompShape','weibLogLik','weibScale','weibShape')
fit = data.frame(matrix(, nrow=length(fit_names), ncol=length(fit_names))) #set up a skeleton table
names(fit) = c( "genotype","media","mat","sd.ls","medianLS","Delta_LL","AvgLS","gompLoglik","gompRate","gompShape","weibLogLik","weibScale","weibShape")


lifespansChar_set<-big_data$set_lifespans
lifespansChar_ref<-big_data$ref_lifespans


```

```{r}
for (genotype_in in c(1:length(big_data$ref_name))){
  
  #genotype=117
  
  
  lifespansChar_set1<-unfactor(lifespansChar_set[genotype_in])
  #if lifespansChar_set1=="NA"
  lifespansTemp_set =as.numeric(unlist(str_split(lifespansChar_set1, ",")))

 
  lifespansChar_ref1<-unfactor(lifespansChar_ref[genotype_in])
  lifespansTemp_ref =as.numeric(unlist(str_split(lifespansChar_ref1, ",")))
  
  lifespansTemp<-c(lifespansTemp_set,lifespansTemp_ref)             
                       
  
  lifespansTemp[lifespansTemp < 0] <- 0
  lifespansTemp <-floor(lifespansTemp+0.5)
  lifespansTemp <- lifespansTemp[ lifespansTemp != 0 ]
  
  lifespanGomp = flexsurvreg(formula = Surv(lifespansTemp) ~ 1, dist = 'gompertz') ### Use the flexsurvreg package to fit lifespan data to gompertz or weibull distribution
  lifespanWeib = flexsurvreg(formula = Surv(lifespansTemp) ~ 1, dist = 'weibull')  
  
## Fill in added columns in the controlConditions table with data from gompertz and weibull ##fitting. Columns are named according to respective variables
  media<-big_data$ref_media
  media=unfactor(media[genotype_in])
  mat<-big_data$ref_mating_type
  mat=unfactor(mat[genotype_in])
  avgLS = mean(lifespansTemp)
  StddevLS = sd(lifespansTemp)
  medianLS = median(lifespansTemp)
  gompShape = lifespanGomp$res[1,1]
  gompRate = lifespanGomp$res[2,1]
  gompLogLik = lifespanGomp$loglik
  gompAIC = lifespanGomp$AIC
  
  weibShape = lifespanWeib$res[1,1]
  weibScale = lifespanWeib$res[2,1]
  weibLogLik = lifespanWeib$loglik
  weibAIC = lifespanWeib$AIC   
  
  delta_LL = lifespanWeib$loglik- lifespanGomp$loglik
  
  genotype<-big_data$ref_name
  
  genotype=unfactor(genotype[genotype_in])
  
#fit_names = c( 'sd.ls','current.seed','i','medianLS','Delta_LL','AvgLS','gompLogLik','gompRat#e','gompShape','weibLogLik','weibScale','weibShape')
  
  fit = rbind(fit,c(genotype,media,mat,StddevLS,medianLS,delta_LL,avgLS,gompLogLik,gompRate,gompShape,weibLogLik,weibScale,weibShape))
  #write.csv(fit, file="Results.csv", row.names=F)
  #names(fit) = c( "genotype","media","mat","sd.ls","medianLS","Delta_LL","AvgLS","gompLoglik","gompRate","gompShape","weibLogLik","weibScale","weibShape")
  
  
#fit= fit[!is.na(fit[,1]), ]
  
}




new_fit<-fit[(length(fit_names)+1):length(fit[,1]),]

write.csv(new_fit,"conditionsWeibRedo_emine_10102016.csv")
```



###read the analyzed data

```{r}
new_fit<-read.csv("conditionsWeibRedo_emine_10102016.csv")

new_fit<-new_fit[-c(1),-c(21)]
Delta_LL<-unlist(new_fit$Delta_LL)
#hist(new_fit$avgLS)



WT.BY4742<- new_fit[new_fit$genotype=="BY4742",]



#WT.BY4742.temp<-WT.BY4742[WT.BY4742$temp==30,]
WT.BY4742.media<-WT.BY4742[WT.BY4742$media=="YPD",]
WT.BY4742.media<-WT.BY4742.media[WT.BY4742.media$mat=="MATalpha",]


#pdf(paste("plots/", "Histogram_empirical_data_WT_BY4742.pdf", sep=''))
par(mfrow=c(3,1)) 
hist(as.numeric(WT.BY4742.media$AvgLS),xlab="Mean Lifespan",main="Wild type: BY4742",
     breaks=length(as.numeric(WT.BY4742.media$AvgLS))/8,col="gray",xlim=c(0,50),ylim=c(0,70))


hist(as.numeric(WT.BY4742.media$gompShape),xlab="G shape parameters",main="",
     breaks=length(as.numeric(WT.BY4742.media$gompShape))/5,col="blue",xlim=c(0,0.16),ylim=c(0,70))

hist(as.numeric(WT.BY4742.media$gompRate),xlab="R rate parameters",main="",
     breaks=length(as.numeric(WT.BY4742.media$gompRate))/6,col="green",xlim=c(0,0.05),ylim=c(0,90))
#dev.off()


WT.BY4741<- new_fit[new_fit$genotype=="BY4741",]



#WT.BY4742.temp<-WT.BY4742[WT.BY4742$temp==30,]
WT.BY4741.media<-WT.BY4741[WT.BY4741$media=="YPD",]

WT.BY4741.media<-WT.BY4741.media[WT.BY4741.media$mat=="MATalpha",]

#pdf(paste("plots/", "Histogram_empirical_data_WT_BY4741.pdf", sep=''))
par(mfrow=c(3,1)) 
hist(as.numeric(WT.BY4741.media$AvgLS),xlab="Mean Lifespan",main="Wild type: BY4741",
     breaks=length(as.numeric(WT.BY4741.media$AvgLS))/5,col="gray",xlim=c(15,40),ylim=c(0,50))


hist(as.numeric(WT.BY4741.media$gompShape),xlab="G shape parameters",main="",
     breaks=length(as.numeric(WT.BY4741.media$gompShape))/5,col="blue",xlim=c(0.05,0.25),ylim=c(0,50))

hist(as.numeric(WT.BY4741.media$gompRate),xlab="R rate parameters",main="",
     breaks=length(as.numeric(WT.BY4741.media$gompRate))/5,col="green",xlim=c(0,0.05),ylim=c(0,70))
#dev.off()


WT.BY4743<- new_fit[new_fit$genotype=="BY4743",]

WT.BY4743.media<-WT.BY4743[WT.BY4743$media=="YPD",]

WT.BY4743.media<-WT.BY4743.media[WT.BY4743.media$mat=="MATalpha",]



#pdf(paste("plots/", "Histogram_empirical_data_WT_BY4743.pdf", sep=''))
par(mfrow=c(3,1)) 
hist(as.numeric(WT.BY4743.media$AvgLS),xlab="Mean Lifespan",main="Wild type: BY4743",
     breaks=length(as.numeric(WT.BY4743.media$AvgLS)),col="gray",xlim=c(25,45),ylim=c(0,30))


hist(as.numeric(WT.BY4743.media$gompShape),xlab="G shape parameters",main="",
     breaks=length(as.numeric(WT.BY4743.media$gompShape)),col="blue",xlim=c(0.05,0.25),ylim=c(0,30))

hist(as.numeric(WT.BY4743.media$gompRate),xlab="R rate parameters",main="",
     breaks=length(as.numeric(WT.BY4743.media$gompRate)),col="green",xlim=c(0,0.01),ylim=c(0,40))
#dev.off()



###check coefficent of the variation: CV
###Because small CV indicate less noisy and more robustness

gompShape<-as.numeric(new_fit$gompShape)
gompRate<-as.numeric(new_fit$gompRate)

CV_gompShape<-sd(gompShape)/mean(gompShape)
CV_gompRate<-sd(gompRate)/mean(gompRate)

sd.Rls<-as.numeric(new_fit$sd.ls)
avg.Rls<-as.numeric(new_fit$AvgLS)


sd.Rls<-sd.Rls[!is.na(sd.Rls)]
avg.Rls<-avg.Rls[!is.na(avg.Rls)]

CV_sd.Rls<- sd(sd.Rls)/mean(sd.Rls)
CV_avg.Rls<-sd(avg.Rls)/mean(avg.Rls)


#How do noises influence likelihood surface of Gompertz and Weibull model? 

WT.BY4742.media$CV<-WT.BY4742.media$sd.ls/WT.BY4742.media$AvgLS

lm( WT.BY4742.media$gompLoglik ~ WT.BY4742.media$CV )

summary(lm( WT.BY4742.media$gompLoglik ~ WT.BY4742.media$CV ))


# We expect that larger CV (more noisy) 
#data will bring down LLH of both Gomeprtz and Weibull models. 
#However, Weibull’s decreasing LLH maybe slower than Gompertz. 


plot(WT.BY4742.media$gompLoglik ~ WT.BY4742.media$CV,col='green',ylab='' )
par(new=TRUE)
plot(WT.BY4742.media$weibLogLik ~ WT.BY4742.media$CV,col='red',ylab='Loglik values')

plot(WT.BY4742.media$gompLoglik ~ WT.BY4742.media$CV )
m<-lm( WT.BY4742.media$gompLoglik ~ WT.BY4742.media$CV )
abline(m,col='red',lty=2,lwd=2)


summary(lm( WT.BY4742.media$weibLogLik ~ WT.BY4742.media$CV))

plot(WT.BY4742.media$weibLogLik ~ WT.BY4742.media$CV )
m<-lm( WT.BY4742.media$weibLogLik ~ WT.BY4742.media$CV )
abline(m,col='red',lty=2,lwd=2)




# We expect that larger CV (more noisy) 
#data will bring down LLH of both Gomeprtz and Weibull models. 
#However, Weibull’s decreasing LLH maybe slower than Gompertz. 



CV_names = c('CV of mean(LS)','CV of sd(LS)', 'CV of gompShape','CV of gompRate')
CV_results = c( CV_avg.Rls,CV_sd.Rls, CV_gompShape,CV_gompRate)
CV.df<-data.frame(CV_names,CV_results)
CV.df

```




