---
title: "Weibull vs Gompertz model comparison with added normal noise"
author: "e guven"
date: "September, 2016"
output: html_document
---
  
  EG started this on 20160403 
how does the noise effect the loglikelihood 
Weibull model->theta is a scale, gamma is a shape parameter >0
Gompertz Model-> G is a scale, R is a shape(rate) parameter >0 

difference function of loglikelihood function of gompertz and weibull p.d.fs
test check how L(Weibull,X)-L(Gompertz,X) values changes for parameters Weibull model->theta is a scale, gamma is a shape parameter >0
Gompertz Model-> G is a scale, R is a shape parameter>0 

For additive Gaussian noise e ~ N (0, sigma^2) with known variance sigma^2
sd of gaussian noise function
max sd would be = 3*mean(inverse.gomp.CDF)
min sd would be mean(inverse.gomp.CDF)

```{r}

require(flexsurv)
require(gplots)

```





```{r}
gamma=0.01
theta=0.25
#test G and R in nested for loops
R=0.001
G=0.2
#R should be in [0, 0.05], and G should be [0.05, 0.3].  
N=2000 # population size

```



```{r}

set.seed(123)
#Gaussian lifespan
normal.random = rnorm(N, mean=50, sd=1)
average.lifespan=mean(normal.random)


lifespan = normal.random
summary(lifespan)

lifespan[lifespan < 0] <- 0
lifespan<-floor(lifespan+0.5)

lifespan <- lifespan[ lifespan != 0 ]




calculate.s = function( lifespan ){
  myData = sort( lifespan[!is.na(lifespan)] );
  tmpGC = table( myData )
  for( i in 2:length(tmpGC)) {
    tmpGC[i] = tmpGC[i-1] + tmpGC[i]        }    
  tot = length(myData)
  tmpGC = tmpGC / tot; 
  s = 1 - tmpGC
  #list( s=s, t=unique(my.data));
  ret = data.frame( cbind(s, unique(myData)));
  names(ret) = c("s", "t");
  ret;
}


GC = calculate.s(lifespan)
plot(GC$s ~ GC$t)



#3)calculate the mortality rates he mortality rate change over time, and then plot log(mortality rate) ~ time.
#Gompertz model should give a linear form.  
#For Weibull model, log(moretality rate) ~ log(time) give the linear form.

#HQin's calculate mortality rate function to calculate the rate of mortality over time
calculate.mortality.rate = function( lifespan ){
  GC = calculate.s(lifespan)
  GC$ds=NA; GC$dt=NA
  #first point
  GC$dt[1] = GC$t[2]
  GC$ds[1] = 1 - GC$s[1]
  GC$mortality.rate[1] = GC$ds[1] / GC$dt[1]
  
  for( j in 2:length(GC[,1])) {
    GC$ds[j] =  GC$s[j-1] - GC$s[j] 
    GC$dt[j] = -GC$t[j-1] + GC$t[j]
    GC$mortality.rate[j] = GC$ds[j] / ( GC$s[j] * GC$dt[j])
  }
  return(GC)
} #end of calculate.mortality.rate()

GC = calculate.mortality.rate(lifespan)



GC = calculate.s(round( lifespan, digits=1))
head(GC)
GC$ds=NA; GC$dt=NA
GC$dt[1] = GC$t[2] #20130321 correct a bug GC$s -> GC$t
GC$ds[1] = 1 - GC$s[1]
GC$mortality.rate[1] = GC$ds[1] / GC$dt[1]

for( j in 2:length(GC[,1])) {
  GC$ds[j] =  GC$s[j-1] - GC$s[j] 
  GC$dt[j] = -GC$t[j-1] + GC$t[j]
  GC$mortality.rate[j] = GC$ds[j] / ( GC$s[j] * GC$dt[j])
}
plot( GC$s ~ GC$t)
plot( GC$mortality.rate ~ GC$t, typ='l', log='y' )

summary(lm(log10(GC$mortality.rate[1:(length(GC$mortality.rate)-1)]) ~ GC$t[1:(length(GC$t)-1)]))
#then plot log(mortality rate) ~ time.
#For Weibull model, log(moretality rate) ~ log(time) give the linear form.

#pdf(paste("plots/","Gompertz.semi.log.plot.batch.pdf", sep=''), width=5, height=5)
plot( log10(GC$mortality.rate) ~ GC$t, type='l') #linear for Gompertz, semi-log plot
text(48,0,"R2= 0.83")
#dev.off()


summary(lm(log10(GC$mortality.rate[1:(length(GC$mortality.rate)-1)]) ~ log10(GC$t[1:(length(GC$t)-1)])))
#pdf(paste("plots/","Weibull.log.log.plot.batch.pdf", sep=''), width=5, height=5)
plot( log10(GC$mortality.rate) ~ log10(GC$t), type='l'  ) #linear for Weibull, log-log plot
text(1.69,0,"R2=0.84")
#dev.off()

```

```{r}

#create a data frame for fit parameters
fit_names = c( 'sd.rnormal','current.seed','scale','sderr','Delta_LL','mean_in','LWei','LGomp','MeanLF','Delta_LL.flex','LWei.flex','LGomp.flex','G.flex.estimated','R.flex.estimated','LLG.par','LLR.par','gamma.flex.estimated','theta.flex.estimated')
fit= data.frame(matrix(, nrow=length(fit_names), ncol=N)) #set up a skeleton table

names(fit) = c( "sd.rnormal","current.seed","scale","sderr","Delta_LL","mean_in","LWei","LGomp","MeanLF","Delta_LL.flex","LWei.flex","LGomp.flex","G.flex.estimated","R.flex.estimated","LLG.par","LLR.par","gamma.flex.estimated","theta.flex.estimated")

```

```{r}
## introduce Weibull model log-likelihood function
#log likelihood function for the Weibull model
#s = exp(-( theta*x)^gamma)
#f = gamma * theta^gamma *x^(gamma-1); 

LL_wei<-function(param,y){
  theta<- exp(param[1]) #take exponential to avoid NaNs when taking log(theta)
  gamma<- exp(param[2]) # avoid NaNs when taking log(gamma)
  
  data=lifespan[!is.na(lifespan)]
  #log_s = -( theta*x)^gamma
  #log_f = log(gamma)+ gamma*log(theta)+ (gamma-1)*log(x) ; 
  #w.lh = sum(log_s)  + sum(log_f);
  
  w.lh<- sum(log(gamma) + gamma*log(theta) + (gamma-1)*log(data)) - sum(
    (theta*data)^gamma)
  
  return(-w.lh)
}
# take log(param) since you take exponential above to avoid NaN values above


##introduce Gompertz model log-likelihood function

#log likelihood function of gompertz model
#s = exp( (R/G) *(1 - exp(G* data)) )  
#f = R * exp( G * data ); 

LL_gomp<- function( param, y ) {
  
  G = exp(param[1]); R = exp(param[2]); 
  data = lifespan[!is.na(lifespan)];
  #log_s = (R/G) *(1 - exp(G* data))
  
  #log_f = log(R) +  G * data ; 
  
  g.lh = sum((R/G) *(1 - exp(G* data)))  + sum(log(R) +  G * data );
  
  return(- g.lh) 
}



## simulate for parameters R, G and scale=i to search effect of noise on delta likl
```



```{r,warnings=F}
for( seed in c(12345, 20160711, -1881, 9999.1234,300045,50758,-10000,74562,-92345,25434)) {
  set.seed(seed)
  current.seed<-seed
  for (mean_in in c(5,10,15,20,35,40,50)){
    
      
      for (i in c(1,2,3,4,5)){ 
        
        
        #generate gompertz random numbers (lifespan) 
        #prediction
        
        normal.random<-rnorm(N, mean=mean_in, sd=i)
        
        lifespan= normal.random
        
        lifespan[lifespan < 0] <- 0
        lifespan<-floor(lifespan+0.5)
        lifespan <- lifespan[ lifespan != 0 ]
        
        summary(lifespan)
        
        scale=i
        sdrnormal<-sd(normal.random)
        
        average.lifespan=mean(lifespan)
        #store average.lifespan into MeanLF list
        MeanLF=average.lifespan
        
        #Log likelihood function for the Weibull model
        
        
        #calculate noise change
        
        sderr = i   
        
        mean_in=mean_in
        
        
        weib=optim(log(c(3,0.03)),LL_wei,y=lifespan)
        LWei = - weib$value
        
        gomp<-optim(param<-log(c(0.01,0.01)), fn=LL_gomp, y=lifespan)
        LGomp= -gomp$value
        
        # store R and G estimation from optim of likl functions in Gompertz
        LLG.par =gomp$par[1]
        LLR.par=gomp$par[2]
        
        delta.likelihood<- -weib$value-(-gomp$value)
        
        
        #Delta_LL[[length(Delta_LL)+1]] = delta.likelihood.wei
        
        
        #flexsurv to calculate the log-likelihood value for both models
        #flexsurv only works with positive variables.
        
        fitGomp = flexsurvreg(formula = Surv(lifespan) ~ 1, dist="gompertz")
        fitWei = flexsurvreg(formula = Surv(lifespan) ~ 1, dist="weibull")
        
        
        LWei.flex=fitWei$loglik
        
        LGomp.flex=fitGomp$loglik
        
        param.Gomp<-fitGomp$res; R.flex<-param.Gomp[2]; G.flex<-param.Gomp[1];
        
        R.flex.estimated<-R.flex
        G.flex.estimated<-G.flex
        
        param.Wei<-fitWei$res; gamma.flex<-param.Wei[1]; theta.flex<-param.Wei[2];
        
        gamma.flex.estimated<-gamma.flex; 
        theta.flex.estimated<-theta.flex
        
        delta_flexsurv=fitWei$loglik-fitGomp$loglik 
        
        #fitWei$loglik
        
        Delta_LL.flex=delta_flexsurv
        
        
        #sim_names = c( "sd.rgompertz","scale","sderr","Delta_LL","G","R","LWei","LGomp","MeanLF","Delta_LL.flex","LWei.flex","LGomp.flex","G.flex.estimated","R.flex.estimated","LLG.par","LLR.par","gamma.flex.estimated","theta.flex.estimated")
        fit = rbind(fit,c(sdrnormal,current.seed,scale,sderr,delta.likelihood,mean_in,LWei,LGomp,MeanLF,Delta_LL.flex,LWei.flex,LGomp.flex,G.flex.estimated,R.flex.estimated,LLG.par,LLR.par,gamma.flex.estimated,theta.flex.estimated))
        #write.csv(fit, file="Results.csv", row.names=F)
        fit= fit[!is.na(fit[,1]), ]
      }
    }
  }

```

fit<- fit[!is.na(names(fit))]
#write.csv(fit, file="Results.normal.csv", row.names=F)

fit<-read.csv(file="Results.normal.csv")


```{r}
summary( lm( fit$Delta_LL~ fit$Delta_LL.flex))
summary( lm( fit$LGomp ~ fit$LGomp.flex))
summary( lm( fit$LWei ~ fit$LWei.flex))

summary(lm(fit$G~fit$G.flex.estimated))
summary(lm(fit$R~fit$R.flex.estimated))


#pdf(paste("plots/","simulated.G.vs.estimated.G.flex_batch.pdf", sep=''), width=5, height=5)
plot(fit$G~fit$G.flex.estimated)
#dev.off()

#pdf(paste("plots/","simulated.R.vs.estimated.R.flex_batch.pdf", sep=''), width=5, height=5)
plot(fit$R~fit$R.flex.estimated)
#dev.off()


#heatmap for fixed G

#fit<-fit[c(1:length(fit[fit$current.seed==12345,5])),]
```

```{r}
Delta_LL=list()
for (i in 1:length(fit[fit$current.seed==12345,5])){
  Delta_LL[length(Delta_LL)+1]<-(fit[fit$current.seed==12345,5][i]+fit[fit$current.seed==20160711,5][i]+
                          fit[fit$current.seed==-1881,5][i]+fit[fit$current.seed==9999.1234,5][i]+fit[fit$current.seed== 300045,5][i]
                        +fit[fit$current.seed==50758,5][i]+fit[fit$current.seed==-10000,5][i]+fit[fit$current.seed== 74562,5][i]+
                          fit[fit$current.seed== -92345,5][i]+fit[fit$current.seed==25434,5][i])/10
  
}




#n.col=256 # number of colors
colfunc<-colorRampPalette(c("aquamarine","cyan","deepskyblue","royalblue","blue"))
#mmx = min(abs(min(fit$Delta_LL)),abs(max(fit$Delta_LL))) # find min value, or you can set to a number
#colbr <- c(seq(-mmx,mmx, len=length(colfunc)+1)) # vector of symmetric breaks

mean.els = unlist(unique(fit$mean_in))
colnum = length(mean.els)

tmp = unlist( unique(fit$scale))
scale.els = tmp[order(tmp)]
rownum = length(scale.els)

mat = matrix( data=NA, nrow= rownum, ncol=colnum) #noise as row, alpha as columns
#rownames(mat) = scale.els
#colnames(mat) = mean.els
rownames(mat) = scale.els
colnames(mat) = mean.els
#for (current.seed in c(12345, 20160711, -1881, 9999.1234,300045,50758,-10000,74562,-92345,25434)){
  
  #data<-fit[fit[,2]==current.seed, 5]
  
  data<-unlist(Delta_LL)
  
  heat_mat<-matrix(data,ncol=colnum,nrow=rownum)
  
  rownames(heat_mat) = scale.els
  colnames(heat_mat) = mean.els
  
  #rownames(heat_mat) <- scale.els
  #colnames(heat_mat) <- mean.els
  
  library(gplots)
  hM <- format(round(heat_mat, 1))
  #data_mat<-scale(heat_mat,scale=TRUE,center=FALSE)

  
  #pdf(paste("plots/", "Gaussian_lifespan_hist_.pdf", sep=''))
  
  hist(heat_mat,xlab=expression(paste( ~delta[LL])),main=expression(paste( ~delta[LL]," of Gaussian lifespans")),
      xlim=c(0,200),ylim=c(0,6),breaks=length(heat_mat),col="blue")
  
  #dev.off()
  #pdf(paste("plots/", "Heatmap_Gaussian.pdf", sep=''))
    
    
    #cellnote=hM,
    heatmap.2(heat_mat,scale="none",col = colfunc, notecol="black",
              dendrogram='none', Rowv=FALSE, Colv=FALSE,trace='none',key=TRUE,symkey=F, density.info="none",
              xlab     = "Mean",
              ylab     = "noise", margins=c(3,3),
                                     main = expression(paste("Heatmap of", ~delta[LL], " for mean vs noise")),par(cex.main=.5),
              srtCol=315, srtRow=315, adjCol = c(0,1),cexRow=0.8,cexCol=0.8,
              sepwidth=c(0.001,0.001),
              rowsep=c(0:length(heat_mat[,1])),colsep=c(0:length(heat_mat[1,])),sepcolor="black")
              
    
 # dev.off()






```










